---
prtitle: A Link Between Poverty and the Shootings of NYC.
author: "Gentaro Nakata"
date: "2023-12-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# A Link Between Poverty and the Shootings of NYC.

## Abstract

In my report, I will be going through data from the New York Police Department that details all the shootings that have happened since 2006. The data set shows what borough the shooting occurred, and the ages and race of the perp and victim. In this report, I will be looking at the data for the number of shootings in each borough of NY, and the race of the perps and victims. We will see that the boroughs with the most shootings were the Bronx and Brooklyn. The data will also show that African Americans are over-represented when it comes to being involved with these crimes. Both communities are known to be some of the poorest in NYC, showing a correlation between poverty and crime.

## Procedure

First, we are going to import the data.

```{r get_jhu_data}
##Data from the NYPD
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
urls <- str_c(url_in)

```{r import_data, message=FALSE}
shooting_cases <- read_csv(urls)
```

Then we are going to filter out the columns with data that we aren't interested in.

```{r tidy_shooting_data}
##Filter out certain columns
shooting_cases <- shooting_cases %>%
               #names_to = "date",
               #values_to = "cases")
  select(-c(LOC_OF_OCCUR_DESC,PRECINCT,JURISDICTION_CODE,LOCATION_DESC,
                         STATISTICAL_MURDER_FLAG,VIC_SEX,X_COORD_CD,Y_COORD_CD,Latitude,Longitude,Lon_Lat,LOC_CLASSFCTN_DESC,PERP_AGE_GROUP,PERP_SEX,VIC_AGE_GROUP))

shooting_cases
```

In the table, the vast majority of the cases that are visible, have had an African-American involved in the incident, whether it's as a perp or a victim.

## Plotting the Data and Analysis

First is the plot for how many incidents happened in which borough of New York City.

```{r borough}
##Plot the Boroughs with the shootings
shooting_cases %>%
  ggplot(aes(x = BORO)) + 
  geom_bar() + 
  labs(title = "Borough of the Crime", x = "Borough", y = "Number of Shootings")
```

The plot shows that the Bronx and Brooklyn have the highest rate of shootings among the boroughs. According to the US Census Data, the Bronx and Brooklyn are the two boroughs with the highest poverty rate in New York. This shows a correlation between high poverty rates and high crime rates.

The second plot is to show the ethnicity of the perp and the victim.

```{r fig.height=10, fig.width=7.5, race, echo=FALSE}
##Plot the race of the perps and that of the victims
shooting_cases %>%
  ggplot(aes(x = VIC_RACE)) + 
  geom_bar() + 
  facet_wrap(vars(PERP_RACE), ncol = 1) + 
  theme(text = element_text(size = 5.5)) +
  labs(title = "Race of Perp and Victim", x = "Race of the Victim", y = "Number of Victims")
```

The plots show that African-Americans are overwhelmingly more likely to be involved with shootings than other ethnicities. Since African-Americans are one of the poorest demographics in NYC, we can once again see the correlation between poverty and crime.

## Biases and Flaws in the Data

There are flaws in this data, for example, there are a large portion of the data where the perp's race isn't specified. Another oversight is that Hispanic people as a whole aren't shown within the data despite them being a large portion of the population of New York in the Census data.
There are also flaws in the way I show the data, I wasn't able to make a graph of the number of shootings that happened each year, thus showing whether the number of shootings in NYC has increased or decreased.

